---
title: "Body Image Trajectories and Parenting — Working Notes"
subtitle: "Analysis notes"
author: "Taren Sanders"
date: today
format:
  docx:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(ggplot2)
library(knitr)

# Quarto renders from the document's directory, so point tar_read() at the
# project-root store explicitly.
store <- here::here("_targets")
```

# Sample overview

```{r overview}
df <- tar_read(df_clean, store = store)
n_obs <- nrow(df)
n_children <- dplyr::n_distinct(df$id)
cohorts <- paste(sort(unique(df$cohort)), collapse = " and ")
ages <- paste(sort(unique(df$age_cat)), collapse = ", ")
```

The analysis dataset contains `r format(n_obs, big.mark = ",")` observations
from `r format(n_children, big.mark = ",")` children across the `r cohorts`
cohorts, measured at ages `r ages`.

::: {.callout-note title="Interpretation"}
- Two cohorts (B and K) are pooled on a common **age** axis (8/10/12).
- Cell sizes are large at every age, so age-stratified estimates are stable.
:::

# Descriptive statistics

```{r descriptives}
descriptives <- tar_read(descriptives_table, store = store)
```

## By cohort within age group

```{r tbl-by-cohort-age}
knitr::kable(
  descriptives$by_cohort_age,
  caption = "Descriptive statistics by cohort within age group."
)
```

::: {.callout-note title="Interpretation"}
- B and K cohorts look very similar at matched ages.
- Suggests that it is reasonable to pool the cohorts for the main analysis.
:::

## By age group

```{r tbl-by-age}
knitr::kable(
  descriptives$by_age,
  caption = "Descriptive statistics by age group (cohorts pooled)."
)
```



::: {.callout-note title="Interpretation"}
- All the means look about what you would expect, and the standard deviations are reasonable.
- The dissatisfaction also looks broadly like you would expect: most children either pick an ideal the same as their current, or a smaller one, with a smaller share picking a larger ideal.
- Body dissatisfaction actually seems to fall slightly with age, counter to what I would have expected (might need to check if this has anything to do with non-response bias?)
- BMI rises with age as expected, but the z-score is consistent. 
- Parental warmth drifts down (but barely) and parental anger is broadly flat across age.
- Parent 1 is clearly the mother in most cases, and parent 2 is the father in most cases, given the match of means.
:::

# Preliminary plots

```{r plot-body-dissat, fig.cap="Body dissatisfaction category proportions by age and cohort.", fig.width=7, fig.height=4}
tar_read(body_dissatisfaction_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- "Larger than ideal" is the dominant form of dissatisfaction at every age.
- The pattern is consistent across the B and K cohorts at matched ages.
:::

# Results

*Placeholder*
