---
title: "Body Image Trajectories and Parenting — Working Notes"
subtitle: "Analysis notes"
author: "Taren Sanders"
date: today
format:
  docx:
    toc: true
    number-sections: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(ggplot2)
library(knitr)

# Quarto renders from the document's directory, so point tar_read() at the
# project-root store explicitly.
store <- here::here("_targets")
```

# Sample overview

```{r overview}
df <- tar_read(df_clean, store = store)
n_obs <- nrow(df)
n_children <- dplyr::n_distinct(df$id)
cohorts <- paste(sort(unique(df$cohort)), collapse = " and ")
ages <- paste(sort(unique(df$age_cat)), collapse = ", ")
```

The analysis dataset contains `r format(n_obs, big.mark = ",")` observations
from `r format(n_children, big.mark = ",")` children across the `r cohorts`
cohorts, measured at ages `r ages`.

::: {.callout-note title="Interpretation"}
- Two cohorts (B and K) are pooled on a common **age** axis (8/10/12).
- Cell sizes are large at every age, so age-stratified estimates are stable.
:::

# Descriptive statistics

```{r descriptives}
descriptives <- tar_read(descriptives_table, store = store)
```

## By cohort within age group

```{r tbl-by-cohort-age}
knitr::kable(
  descriptives$by_cohort_age,
  caption = "Descriptive statistics by cohort within age group."
)
```

::: {.callout-note title="Interpretation"}
- B and K cohorts look very similar at matched ages.
- Suggests that it is reasonable to pool the cohorts for the main analysis.
:::

## By age group

```{r tbl-by-age}
knitr::kable(
  descriptives$by_age,
  caption = "Descriptive statistics by age group (cohorts pooled)."
)
```



::: {.callout-note title="Interpretation"}
- All the means look about what you would expect, and the standard deviations are reasonable.
- The direction-agnostic dissatisfaction indicator ("Any Direction") shows just under half of children pick an ideal different from their current figure at every age.
- The dissatisfaction also looks broadly like you would expect: most children either pick an ideal the same as their current, or a smaller one, with a smaller share picking a larger ideal.
- Body dissatisfaction actually seems to fall slightly with age, counter to what I would have expected (might need to check if this has anything to do with non-response bias?)
- BMI rises with age as expected, but the z-score is consistent. 
- Parental warmth drifts down (but barely) and parental anger is broadly flat across age.
- Parent 1 is clearly the mother in most cases, and parent 2 is the father in most cases, given the match of means.
:::

## By sex within age group

```{r tbl-by-sex-age}
knitr::kable(
  descriptives$by_sex_age,
  caption = "Descriptive statistics by sex within age group (cohorts pooled)."
)
```

::: {.callout-note title="Interpretation"}
- The decline in dissatisfaction with age noted above is mostly a boys' pattern: any-direction dissatisfaction falls steadily for boys (53.5% → 44.5% → 41.1%), while for girls it drops from 8 to 10 (52.2% → 44.8%) and then rises again at 12 (47.2%).
- The rebound for girls at 12 is driven by "larger than ideal" (38.9% → 42.6%), whereas for boys "larger than ideal" keeps falling (40.3% → 36.0% → 32.1%) — plausibly early-puberty timing for girls.
- "Thinner than ideal" declines with age in both sexes but is roughly twice as common in boys at every age.
- Together this suggests both the level (at older ages) and especially the direction of dissatisfaction are sex-related, so models should be sex-stratified.

:::

# Preliminary plots

```{r plot-body-dissat, fig.cap="Body dissatisfaction category proportions by age and cohort.", fig.width=7, fig.height=4}
tar_read(body_dissatisfaction_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- "Larger than ideal" is the dominant form of dissatisfaction at every age.
- The pattern is consistent across the B and K cohorts at matched ages.
:::

# Analysis sample

```{r sample-flow}
sample_flow <- tar_read(sample_flow_table, store = store)
analysis_n <- sample_flow$flow$`N children`[3]
regression_n <- sample_flow$flow$`N children`[5]
sensitivity_n <- sample_flow$flow$`N children`[6]
```

The trajectory analyses use children with a CBIS body dissatisfaction score at
two or more of the three ages (N = `r format(analysis_n, big.mark = ",")`).
The primary class-membership regressions additionally require complete baseline
(age 8) covariates — sex, SES, BMI z-score, and Parent 1 warmth and anger
(N = `r format(regression_n, big.mark = ",")`). A sensitivity model adding
Parent 2 warmth and anger uses the subset with second-carer data
(N = `r format(sensitivity_n, big.mark = ",")`).

```{r tbl-sample-flow}
knitr::kable(
  sample_flow$flow,
  caption = "Sample flow from the cleaned data to the analysis samples."
)
```

```{r tbl-baseline-missingness}
knitr::kable(
  sample_flow$baseline_missingness,
  caption = "Baseline (age 8) covariate missingness in the trajectory sample."
)
```

::: {.callout-note title="Interpretation"}
- Attrition on the outcome is modest: 91% of children contribute at least two
  CBIS waves, and 75% contribute all three, so the trajectory sample retains
  most of the cohort.
- Parent 2 warmth/anger are missing for ~38% of the trajectory sample —
  largely single-parent households or no second-carer questionnaire — while
  every other covariate is missing for under 9%. Requiring Parent 2 scores
  would cost ~2,300 children.
- **Decision (2026-07-23):** the primary regressions therefore use Parent 1
  parenting only, keeping the larger sample; the both-parent model (all four
  parenting variables jointly, as in the original proposal) becomes a
  sensitivity analysis on the subset with second-carer data. Note the
  sensitivity subset is also selectively two-parent households, so any
  differences may reflect the sample as much as mutual adjustment.
- The 129 children "missing" sex are those with no age-8 record at all (their
  baselines are taken from the age-8 wave); they are excluded from the
  regressions anyway by their other missing baselines.
:::

# Trajectory class enumeration

```{r lcga-fit-stats}
lcga_stats <- tar_read(lcga_fit_stats, store = store)
lcga_display <- lcga_stats |>
  dplyr::transmute(
    K = k,
    Converged = ifelse(converged, "Yes", "No"),
    `Log-likelihood` = sprintf("%.1f", loglik),
    Parameters = npm,
    AIC = sprintf("%.0f", AIC),
    BIC = sprintf("%.0f", BIC),
    SABIC = sprintf("%.0f", SABIC),
    Entropy = ifelse(is.na(entropy), "—", sprintf("%.2f", entropy)),
    `Class % (largest → smallest)` = class_props
  )
```

Latent class growth analyses (linear class trajectories, no within-class
random effects, equal residual variance across classes) were fit for K = 1–5
classes with `lcmm::hlme()`, each multi-class model from 50 random starts
(gridsearch initialised from the one-class model).

```{r tbl-lcga-fit}
knitr::kable(
  lcga_display,
  caption = "Fit statistics for K = 1-5 latent class growth models of body dissatisfaction.",
  align = c("l", rep("r", ncol(lcga_display) - 1))
)
```

```{r plot-lcga-trajectories, fig.cap="Predicted class trajectories (95% bands) by number of classes. Dashed line: observed overall means.", fig.width=10, fig.height=3.5}
tar_read(lcga_trajectory_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- All five models converged. As anticipated at this sample size (N = 7,843),
  AIC/BIC/SABIC fall monotonically through K = 5, so information criteria
  alone cannot settle enumeration — entropy, class sizes, and
  interpretability carry more weight.
- **Entropy peaks at K = 3 (0.85)** and drops to ~0.81 at K = 4–5. All values
  from K = 3 up clear the ≥0.8 bar for the proportional-weighting three-step
  approach (Step 4 decision point).
- The K = 3 solution is highly interpretable: a satisfied majority tracking
  close to zero discrepancy (~84%), a "larger than ideal" class starting high
  and declining (~13.5%), and a small "thinner than ideal" class converging
  toward zero (~2.8%, ≈ 220 children).
- K = 4 splits the "larger than ideal" group into a moderate-stable class
  (~23%) and a small extreme-declining class (~2.7%), at the cost of lower
  entropy; K = 5 adds a 1.6% class with a steep crossing trajectory that looks
  like a fringe pattern rather than a substantive group.
- Classes below the 5% flag appear from K = 3 onward, but the 2.8% class at
  K = 3 maps directly onto the "thinner than ideal" group visible in the
  descriptives, which argues for keeping it.
- **Tentative recommendation: K = 3** (best entropy, cleanly interpretable,
  classes matching the descriptive patterns), with K = 4 the main alternative
  if the moderate/extreme split among dissatisfied children is of substantive
  interest. **Decision on K is yours** — it is set in one place
  (`lcga_k_chosen` in `_targets.R`) at Step 3b.
:::

## Sex-stratified enumeration check

The descriptive statistics showed sex-linked patterns in the level and
direction of body dissatisfaction, raising the question of whether trajectory
classes should be sex-stratified. As a diagnostic, the same K = 1–5
enumeration was repeated separately for boys and girls. If the stratified
class *shapes* mirror the pooled solution, sex differences are a matter of
class *prevalence* — handled by sex in the class-membership model and the
planned sex-moderation tests — and the pooled LCGA stands as the primary
specification.

```{r lcga-strat-stats}
lcga_strat_stats <- tar_read(lcga_strat_fit_stats, store = store)
strat_display <- dplyr::bind_rows(lcga_stats, lcga_strat_stats) |>
  dplyr::arrange(sample, k) |>
  dplyr::transmute(
    Sample = tools::toTitleCase(sample),
    K = k,
    Converged = ifelse(converged, "Yes", "No"),
    BIC = sprintf("%.0f", BIC),
    SABIC = sprintf("%.0f", SABIC),
    Entropy = ifelse(is.na(entropy), "—", sprintf("%.2f", entropy)),
    `Class % (largest → smallest)` = class_props
  )
```

```{r tbl-lcga-strat}
knitr::kable(
  strat_display,
  caption = "Class enumeration in the pooled sample and stratified by sex.",
  align = c("l", rep("r", ncol(strat_display) - 1))
)
```

```{r plot-lcga-sex-comparison, fig.cap="Predicted class trajectories by sample (pooled, boys, girls) and number of classes. Dashed lines: observed means within each sample.", fig.width=10, fig.height=8}
tar_read(lcga_sex_comparison_plot, store = store)
```

::: {.callout-note title="Interpretation"}
- All stratified models converged, and enumeration behaves the same way
  within each sex as in the pooled sample: BIC falls monotonically, entropy
  peaks at K = 3 (0.85 in boys, girls, and pooled alike).
- **At K = 3 the class shapes replicate almost exactly in both sexes**: a
  flat satisfied majority (~84% in both), a declining "larger than ideal"
  class (boys 12.6%, girls 14.3%), and a rising "thinner than ideal" class
  (boys 3.5%, girls 2.0%). The sexes differ in class *prevalence* — the
  thinner-than-ideal class is nearly twice as common in boys — not in class
  *shape*. Prevalence differences are exactly what the class-membership
  model with sex handles.
- The girls' age-12 rebound in "larger than ideal" seen in the descriptives
  surfaces only at K = 4, and only as a slope difference: girls' moderate
  dissatisfied class drifts slightly upward from 10 to 12 while the boys'
  equivalent drifts down. It is a modest divergence within one class, not a
  distinct trajectory structure, and no rising-positive class exists in the
  girls-only solutions at K = 3.
- **Conclusion: the pooled LCGA stands as the primary specification.** This
  supports K = 3 specifically — its structure is reproducible within each
  sex, whereas the K = 4 structures start to diverge across sexes. If the
  moderate-class slope difference matters substantively, it is better
  captured by the planned sex-moderation analysis than by stratified
  classes.
:::

::: {.callout-important title="Enumeration criteria & limitations"}
- Class enumeration rests on BIC/SABIC, entropy, smallest-class size (flag
  <5%), and interpretability. `lcmm` provides no LMR-LRT or BLRT (Nylund et
  al., 2007); a manual bootstrap LRT is a possible add-on if the picture is
  ambiguous.
- At this sample size BIC can keep falling as classes are added — where it
  does, we lean on entropy, class sizes, and interpretability.
- Never accept a solution with `Converged = No`.
- Linear class trajectories only: three timepoints saturate a quadratic.
- Reporting follows the GRoLTS checklist (van de Schoot et al., 2017).
:::

# Trajectory classes (final model)

```{r final-classes}
class_assignments <- tar_read(class_assignments, store = store)
class_quality <- class_assignments |>
  dplyr::count(class, name = "N") |>
  dplyr::left_join(
    class_assignments |>
      dplyr::group_by(class) |>
      dplyr::summarise(
        mean_pp = mean(modal_pprob),
        min_pp = min(modal_pprob)
      ),
    by = "class"
  ) |>
  dplyr::transmute(
    Class = class,
    N,
    `%` = sprintf("%.1f", 100 * N / sum(N)),
    `Mean posterior prob.` = sprintf("%.2f", mean_pp),
    `Min posterior prob.` = sprintf("%.2f", min_pp)
  )
```

The three-class model was frozen as the final trajectory model (K = 3;
entropy 0.85), based on the enumeration statistics and the sex-stratified
check above. Classes are labelled in ascending order of their age-8
intercept, so numbering is stable across refits.

```{r plot-lcga-classes, fig.cap="Final three-class model: predicted trajectories (95% bands) and observed means by modal class.", fig.width=7, fig.height=4.5}
tar_read(lcga_class_plot, store = store)
```

```{r tbl-class-quality}
knitr::kable(
  class_quality,
  caption = "Class sizes and modal-assignment quality (posterior probabilities of the assigned class).",
  align = c("l", rep("r", 4))
)
```

::: {.callout-note title="Interpretation"}
- **Class 1 — "thinner than ideal, converging" (2.8%, n = 218):** starts
  around −1.7 at age 8 and rises toward −0.4 by 12; these children see
  themselves as thinner than their ideal, with the gap closing over time.
- **Class 2 — "satisfied" (83.8%, n = 6,570):** flat trajectory just above
  zero across ages 8–12.
- **Class 3 — "larger than ideal, declining" (13.5%, n = 1,055):** starts
  around 1.7 and declines modestly to about 1.3; persistent dissatisfaction
  in the "feels larger than ideal" direction.
- Classification quality is good: mean posterior probabilities of the
  assigned class are 0.81 / 0.94 / 0.90 (all above the 0.70 rule of thumb),
  consistent with entropy of 0.85. Observed class means track the model
  predictions closely, with mild shrinkage at the age-8 extremes.
- Entropy ≥ 0.8 settles the Step 4 decision point: the three-step membership
  models will use **proportional posterior weighting**, with a
  modal-assignment sensitivity check.
- Class names above are working labels — happy to change them to whatever
  you want used in the manuscript.
:::

# Results

*Placeholder*
